---
title: "Mouse Lingon diet experiment: compare with the published results"
author:
   name: "Shuyi Li & Dmytro Kryvokhyzha"
   email: dmytro.kryvokhyzha@med.lu.se
   affiliation: LUDC Bioinformatics Unit
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
    html_document:
      toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, eval=TRUE, message=FALSE)
knitr::opts_knit$set(root.dir = '../')
```

## Data

```{r}
library(dplyr)
library(VennDiagram)
```

DESeq_res: the DESeq results from the musLingon project

linear_res: the ANOVA results from the insulin signaling project

```{r}
DEseq_res <- read.table("results/tables/deseq/LingonProj_DESeqres.csv", header = TRUE)
linear_res <- read.csv("data/paper/table_linear_expression_pvals.20140702.csv",header = TRUE)
anova_res <- read.table("data/paper/FAT_table_anova_Fstat_expression_pvals.20140707.csv",
                        header = TRUE, sep = "\t", quote = "\"",)
```

## Correlation in all genes

### Mean expression levels

Calculate the means

```{r}
DEseq_res_mean <- DEseq_res[,c('gene','HFD_tpm','LFD_tpm')]
linear_res_mean <- data.frame(gene = linear_res$ensembl_gene_id, 
                              fat_D0_mean = rowMeans(linear_res[,
                                 grep("D0", colnames(linear_res))]),
                              fat_D2_mean = rowMeans(linear_res[,
                                 grep("D2", colnames(linear_res))]),
                              fat_D4_mean = rowMeans(linear_res[,
                                 grep("D4", colnames(linear_res))]),
                              fat_D6_mean = rowMeans(linear_res[,
                                 grep("D6", colnames(linear_res))]),
                              fat_D14_mean = rowMeans(linear_res[,
                                 grep("D14", colnames(linear_res))]))

## select common DE genes
common_expr_gene <- inner_join(DEseq_res_mean, linear_res_mean, by = 'gene')
```

Correlation

```{r}
knitr::kable(cor(as.matrix(common_expr_gene[,-1]), method ="spearman"))
```

### Fold change

Get fold changes

```{r}
anova_factor <- anova_res[,c(1, grep("factor", colnames(anova_res)))]
colnames(anova_factor)[1] <- 'gene'
DEseq_res_fold <- DEseq_res[,c('gene','log2FoldChange_HFD_vs_LFD')]
DESeq_anova_fold <- inner_join(DEseq_res_fold, anova_factor, by = 'gene')

linear_beta <- linear_res[,c(1, grep("Beta", colnames(linear_res)))]
colnames(linear_beta)[1] <- 'gene'
DESeq_linear_fold <- inner_join(DEseq_res_fold, linear_beta, by = 'gene')
```

#### Linear regression

```{r}
knitr::kable(cor(na.omit(as.matrix(DESeq_linear_fold[,-1])), method ="spearman"))
```

#### ANOVA

```{r}
knitr::kable(cor(na.omit(as.matrix(DESeq_anova_fold[,-1])), method ="spearman"))
```

## Correlation in significant

Get the DE genes in all results

```{r}
DEseq_sign <- DEseq_res$gene[which(DEseq_res$padj_HFD_vs_LFD <= 0.01)]
linear_sign <- linear_res$ensembl_gene_id[which(linear_res$adj.P.Val <= 0.1)]
anova_sign <- anova_res$ensembl_gene_id[which(anova_res$adj.P.Val <= 0.01)]
length(DEseq_sign)
length(linear_sign)
length(anova_sign)

DEseq_linear_sign <- union(DEseq_sign, linear_sign)
DEseq_anova_sign <- union(DEseq_sign, anova_sign)
length(DEseq_linear_sign)
length(DEseq_anova_sign)
```

### Mean expression levels

Get the means

```{r}
DEseq_linear_sign_means <- common_expr_gene[which(common_expr_gene$gene %in% DEseq_linear_sign),]
DEseq_anova_sign_means <- common_expr_gene[which(common_expr_gene$gene %in% DEseq_anova_sign),]
```

#### Linear regression

```{r}
knitr::kable(cor(as.matrix(DEseq_linear_sign_means[,-1]), method ="spearman"))
```

#### ANOVA

```{r}
knitr::kable(cor(as.matrix(DEseq_anova_sign_means[,-1]), method ="spearman"))
```

### Fold change

Get fold changes

```{r}
DESeq_linear_fold_sign <- DESeq_anova_fold[which(DESeq_linear_fold$gene %in% DEseq_linear_sign),]
DESeq_anova_fold_sign <- DESeq_anova_fold[which(DESeq_anova_fold$gene %in% DEseq_anova_sign),]
```

#### Linear regression

```{r}
knitr::kable(cor(na.omit(as.matrix(DESeq_linear_fold_sign[,-1])), method ="spearman"))
```

#### ANOVA

```{r}
knitr::kable(cor(na.omit(as.matrix(DESeq_anova_fold_sign[,-1])), method ="spearman"))
```

## Correlation without lowly expressed genes

Remove the genes with the expression lower than 80% genes

```{r}
DEseq_res_fil <- DEseq_res$gene[order(DEseq_res$HFD_tpm, decreasing=TRUE)
                                ][1:(length(DEseq_res$HFD_tpm)*0.8)]
linear_res_fil <- linear_res_mean$gene[order(linear_res_mean$fat_D4_mean, decreasing=TRUE)
                                       ][1:(length(linear_res_mean$fat_D4_mean)*0.8)]
common_expr_gene_fil <- intersect(DEseq_res_fil, linear_res_fil)
```

### Mean expression levels

Get the means

```{r}
common_expr_mean_fil <- common_expr_gene[which(common_expr_gene$gene %in% common_expr_gene_fil),]
```

#### Linear regression

```{r}
knitr::kable(cor(as.matrix(common_expr_mean_fil[,-1]), method ="spearman"))
```

### Fold change

Get fold changes

```{r}
DESeq_linear_fold_fil <- DESeq_linear_fold[which(DESeq_linear_fold$gene %in%
                                                    common_expr_gene_fil),]
DESeq_anova_fold_fil <- DESeq_anova_fold[which(DESeq_anova_fold$gene %in%
                                                  common_expr_gene_fil),]
```

#### Linear regression

```{r}
knitr::kable(cor(na.omit(as.matrix(DESeq_linear_fold_fil[,-1])), method ="spearman"))
```

#### ANOVA

```{r}
knitr::kable(cor(na.omit(as.matrix(DESeq_anova_fold_fil[,-1])), method ="spearman"))
```

## Correlation in top 2000 most expressed genes

Get the top 2000 genes in all results

```{r}

DEseq_2000 <- DEseq_res$gene[order(DEseq_res$HFD_tpm, decreasing=TRUE)][c(1:2000)]
linear_2000 <- linear_res_mean$gene[order(linear_res_mean$fat_D4_mean, decreasing=TRUE)][c(1:2000)]
common_expr_gene_2000 <- intersect(DEseq_2000, linear_2000)
```

### Mean expression levels

Get the means

```{r}
common_expr_mean_2000 <- common_expr_gene[which(common_expr_gene$gene %in% common_expr_gene_2000),]
```

#### Linear regression

```{r}
knitr::kable(cor(as.matrix(common_expr_mean_2000[,-1]), method ="spearman"))
```

### Fold change

Get fold changes

```{r}
DESeq_linear_fold_2000 <- DESeq_linear_fold[which(DESeq_linear_fold$gene %in%
                                                    common_expr_gene_2000),]
DESeq_anova_fold_2000 <- DESeq_anova_fold[which(DESeq_anova_fold$gene %in%
                                                  common_expr_gene_2000),]
```

#### Linear regression

```{r}
knitr::kable(cor(na.omit(as.matrix(DESeq_linear_fold_2000[,-1])), method ="spearman"))
```

#### ANOVA

```{r}
knitr::kable(cor(na.omit(as.matrix(DESeq_anova_fold_2000[,-1])), method ="spearman"))
```

### Overlap in the DE genes in both project

Select significant genes

```{r}
DEseq_sign_venn <- DEseq_res$gene[which(DEseq_res$padj_HFD_vs_LFD <= 0.05)]
linear_sign_venn <- linear_res$ensembl_gene_id[which(linear_res$adj.P.Val <= 0.1)]
anova_sign_venn <- anova_res$ensembl_gene_id[which(anova_res$adj.P.Val <= 0.05)]
```

All three comparisons

```{r}
comparisons_genes <- list(DEseq_sign_venn,
                         linear_sign_venn,
                         anova_sign_venn)

comparisons_names <- c('DESeq (current)',
                       'Linear regression (14 days paper)',
                       'ANOVA (14 days paper)')

vennPlot <- venn.diagram(comparisons_genes,
                        NULL,
                        fill = c("red", "blue", "green"),
                        category.names = comparisons_names,
                        cex=1.5,
                        main.cex = 1.5,
                        sub.cex = 1.5)
grid.draw(vennPlot)
```

DESeq vs ANOVA comparisons

```{r}
comparisons_genes2 <- list(DEseq_sign_venn, anova_sign_venn)

comparisons_names2 <- c('DESeq',
                       'ANOVA')

vennPlot2 <- venn.diagram(comparisons_genes2,
                        NULL,
                        fill = c("red", "blue"),
                        category.names = comparisons_names2,
                        cex=1.5,
                        main.cex = 1.5,
                        sub.cex = 1.5)
grid.draw(vennPlot2)
```

```{bash clean_logs, include=FALSE}
rm VennDiagram*.log
```
