---
title: "LingonProj Gene Comparison"
author:
   name: "Shuyi Li"
   email: shuyi.li@med.lu.se
   affiliation: LUDC Bioinformatics Unit
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, eval=TRUE, message=FALSE)
knitr::opts_knit$set(root.dir = '../')
```

## Load data
```{r}
library(dplyr)
```

DESeq_res: the DESeq results from the musLingon project

linear_expression: the ANOVA results from the insulin signaling project

```{r}
DEseq_res <- read.table("data/LingonProj_DESeqres.csv", header = TRUE)
linear_expression <- read.csv("data/table_linear_expression_pvals.20140702.csv",header = TRUE)
anova_factors <- read.table("data/anova_factors.csv", header = TRUE)
```

## Create new matrix for comparison

calculate the mean expression level

```{r}
DEseq_res_mean <- DEseq_res[,c('gene','HFD_tpm','LFD_tpm')]
linear_expression_mean<- data.frame(gene = linear_expression$ensembl_gene_id, 
                                    fat_D0_mean = rowMeans(linear_expression[, c(11:14)]),
                                    fat_D2_mean = rowMeans(linear_expression[, c(15:18)]),
                                    fat_D4_mean = rowMeans(linear_expression[, c(19:22)]),
                                    fat_D6_mean = rowMeans(linear_expression[, c(23:26)]),
                                    fat_D14_mean = rowMeans(linear_expression[, c(27:30)]))

## select common DE genes
common_expr_gene <- inner_join(DEseq_res_mean, linear_expression_mean, by = 'gene')
head(common_expr_gene)

names(anova_factors) <- c('gene','Day.2','Day.4','Day.6','Day.14')
DEseq_res_fold <- DEseq_res[,c('gene','log2FoldChange_HFD_vs_LFD')]
common_fold <- inner_join(DEseq_res_fold, anova_factors, by = 'gene')
```

## Correlation

### correlation between expression level

plot the data

```{r}
plot(common_expr_gene[, 4],log2(common_expr_gene[, 2]),xlab="fat_D0_mean", ylab="HFD_tpm")
plot(common_expr_gene[, 5],log2(common_expr_gene[, 2]),xlab="fat_D2_mean", ylab="HFD_tpm")
plot(common_expr_gene[, 6],log2(common_expr_gene[, 2]),xlab="fat_D4_mean", ylab="HFD_tpm")
plot(common_expr_gene[, 7],log2(common_expr_gene[, 2]),xlab="fat_D6_mean", ylab="HFD_tpm")
plot(common_expr_gene[, 8],log2(common_expr_gene[, 2]),xlab="fat_D14_mean", ylab="HFD_tpm")

plot(common_expr_gene[, 4],log2(common_expr_gene[, 2]),xlab="fat_D0_mean", ylab="LFD_tpm")
plot(common_expr_gene[, 5],log2(common_expr_gene[, 2]),xlab="fat_D2_mean", ylab="LFD_tpm")
plot(common_expr_gene[, 6],log2(common_expr_gene[, 2]),xlab="fat_D4_mean", ylab="LFD_tpm")
plot(common_expr_gene[, 7],log2(common_expr_gene[, 2]),xlab="fat_D6_mean", ylab="LFD_tpm")
plot(common_expr_gene[, 8],log2(common_expr_gene[, 2]),xlab="fat_D14_mean", ylab="LFD_tpm")

```


```{r}
cor_expr<- matrix(0,7,7, dimnames = list(colnames(common_expr_gene)[2:8], colnames(common_expr_gene)[2:8]))

for(i in 2:8){
   for(j in 2:8){
     t <- cor.test(common_expr_gene[, i], common_expr_gene[, j], method ="spearman", exact = FALSE)
     cor_expr[i-1, j-1] <- t$estimate
   }
}
print(cor_expr)
```

### correlation between fold change

plot the data

```{r}
plot(common_fold[, 3],common_fold[, 2],xlab="D2.factor", ylab="log2FoldChange_HFD_vs_LFD")
plot(common_fold[, 4],common_fold[, 2],xlab="D4.factor", ylab="log2FoldChange_HFD_vs_LFD")
plot(common_fold[, 5],common_fold[, 2],xlab="D6.factor", ylab="log2FoldChange_HFD_vs_LFD")
plot(common_fold[, 6],common_fold[, 2],xlab="D14.factor", ylab="log2FoldChange_HFD_vs_LFD")

```


```{r}
cor_fold<- matrix(0,5,5, dimnames = list(colnames(common_fold)[2:6], colnames(common_fold)[2:6]))

for(i in 2:6){
   for(j in 2:6){
     t <- cor.test(common_fold[, i], common_fold[, j], method ="spearman", exact = FALSE)
     cor_fold[i-1, j-1] <- t$estimate
   }
}
print(cor_fold)
```



## take the overlapped significant DE genes in both project

```{r}
DEseq_res_sign <- DEseq_res[DEseq_res$padj_HFD_vs_LFD < 0.05,]
anova_res_sign <- linear_expression[linear_expression$P.Val < 0.05,]

DEseq_res_sign <- DEseq_res[,c('gene','HFD_tpm','LFD_tpm')]
anova_res_sign<- data.frame(gene = anova_res_sign$ensembl_gene_id, 
                                    fat_D0_mean = rowMeans(anova_res_sign[, c(11:14)]),
                                    fat_D2_mean = rowMeans(anova_res_sign[, c(15:18)]),
                                    fat_D4_mean = rowMeans(anova_res_sign[, c(19:22)]),
                                    fat_D6_mean = rowMeans(anova_res_sign[, c(23:26)]),
                                    fat_D14_mean = rowMeans(anova_res_sign[, c(27:30)]))

common_sign <- inner_join(DEseq_res_sign, anova_res_sign, by = 'gene')

```

### correlation between expression level

```{r}
cor_expr_sign<- matrix(0,7,7, dimnames = list(colnames(common_sign)[2:8], colnames(common_sign)[2:8]))

for(i in 2:8){
   for(j in 2:8){
     t <- cor.test(common_sign[, i], common_sign[, j], method ="spearman", exact = FALSE)
     cor_expr_sign[i-1, j-1] <- t$estimate
   }
}
print(cor_expr_sign)
```

## remove low expressed genes (the expression level of the gene is lower than 80% genes)

```{r}
DEseq_res_fil <- DEseq_res[order(DEseq_res$HFD_tpm, decreasing=TRUE), ][1:(54347*0.8), ]
DEseq_res_fil <- DEseq_res_fil[,c('gene','HFD_tpm','LFD_tpm')]
linear_expression_fil <- linear_expression_mean[order(linear_expression_mean$fat_D4_mean, decreasing=TRUE), ][1:(39179*0.8), ]
common_expr_gene_fil <- inner_join(DEseq_res_fil, linear_expression_fil, by = 'gene')

```

### correlation between expression level

```{r}
cor_expr_fil<- matrix(0,7,7, dimnames = list(colnames(common_expr_gene_fil)[2:8], colnames(common_expr_gene_fil)[2:8]))

for(i in 2:8){
   for(j in 2:8){
     t <- cor.test(common_expr_gene_fil[, i], common_expr_gene_fil[, j], method ="spearman", exact = FALSE)
     cor_expr_fil[i-1, j-1] <- t$estimate
   }
}
print(cor_expr_fil)
```

## take the top 2000 expressed genes

```{r}
DEseq_res_top2000 <- DEseq_res[order(DEseq_res$HFD_tpm, decreasing=TRUE), ][1:2000, ]
DEseq_res_top2000  <- DEseq_res_top2000[,c('gene','HFD_tpm','LFD_tpm')]
linear_expression_top2000 <- linear_expression_mean[order(linear_expression_mean$fat_D4_mean, decreasing=TRUE), ][1:2000, ]
common_expr_gene_top <- inner_join(DEseq_res_top2000, linear_expression_top2000, by = 'gene')

```

### correlation between expression level

```{r}
cor_expr_top<- matrix(0,7,7, dimnames = list(colnames(common_expr_gene_top)[2:8], colnames(common_expr_gene_top)[2:8]))

for(i in 2:8){
   for(j in 2:8){
     t <- cor.test(common_expr_gene_top[, i], common_expr_gene_top[, j], method ="spearman", exact = FALSE)
     cor_expr_top[i-1, j-1] <- t$estimate
   }
}
print(cor_expr_top)
```
